How Optimal is Word Recognition Under Multimodal Uncertainty?

### 2019-04-12 Writeup/ms.tex
more revisions

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={How Optimal is Novel Word Recogntion Under Multimodal Uncertainty?},
            pdfauthor={Abdellah Fourtassi~\& Michael C. Frank},
            pdfkeywords={Language understanding; audio-visual processing; word learning; speech
perception; computational modeling.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{How Optimal is Novel Word Recogntion Under Multimodal Uncertainty?}
    \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C.
Frank\textsuperscript{1}}
    \date{}

@@

\abstract{
Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We introduce
a paradigm that allows us to examine how adults identify novel words
under joint uncertainty in the auditory and visual modalities and
propose an ideal observer model of how cues in these modalities are
combined optimally. Model predictions are tested in three experiments
where novel word recognition is made under two kinds of uncertainty:
category ambiguity and perceptual noise. In all cases, the optimal model
explains much of the variance in human mean judgments. When the signal
is not distorted with noise, participants weight the auditory and visual
cues optimally, that is, according to the relative reliability of each
modality. But when one modality has noise added to it, human perceivers
systematically prefer the unperturbed modality to a greater extent than
the optimal model does. The study provides a formal framework which
helps to quantify how word form and word meaning interact in word
recognition under uncertainty. Moreover it offers a first step towards a
model that accounts for form-meaning synergies in early word learning.

@@

Language uses symbols expressed in one modality -- the auditory
modality, in the case of speech -- to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee.} Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

As a simplified case study of early word learning, the task of matching
sounds to corresponding visual objects has been studied extensively in
the developmental literature. For example, many studies focus on how
children might succeed in this type of task despite referential
ambiguity (Medina, Snedeker, Trueswell, \& Gleitman, 2011; Pinker, 1989;
Smith \& Yu, 2008; Suanda, Mugwanya, \& Namy, 2014; Vlach \& Johnson,
2013; Vouloumanos, 2008; Yurovsky \& Frank, 2015). However, even when
they \emph{know} the meanings of a word, listeners (both children and
adults) often still find it challenging to recognize which word the
speaker has uttered, especially under noise (Mattys, Davis, Bradlow, \&
Scott, 2012; Peelle, 2018). The purpose of the current study is thus to
explore word recognition by adults under multimodal uncertainty. We
focus on the special case where people have access to multimodal cues
from the auditory speech and the visual referent. In the General
Discussion, we return to the question of how these findings relate to
questions about word learning.

@@

Some prior ideal observer studies are closely related to the question we
are addressing in the current work. For instance, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (Voice Onset Time) that differentiated similar
words (e.g., \enquote{beach} and \enquote{peach}). They found that
humans were sensitive to these probabilistic cues and their judgments
closely reflected the optimal predictions. And Feldman et al. (2009)
studied the perceptual magnet effect, a phenomenon that involves reduced
discriminability near prototypical sounds in the native language (Kuhl,
1991), showing that this effect can be explained as the consequence of
optimally solving the problem of perception under uncertainty.

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the predictions of an ideal observer. This
previous research did not, however, systematically study speech
understanding when visual information was obtained through the
referential context rather than through observation of speaker's face.
Although some experimental findings show that information about the
identity of a referent can be integrated with linguistic information to
resolve lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995), to our knowledge no study has offered an ideal observer
analysis of this task.

Combining information between words and visual referents might seem
similar to audio-visual speech integration, but there are at least two
fundamental differences between these two cases, and both can influence
the way the auditory and visual cues are combined.

First, in the case of audio-visual speech, both modalities offer
information about the same underlying speech category. They differ only
in terms of their informational reliability. In a referential context,
however, the auditory and visual modalities play different roles in the
referential process -- the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning} (and these
differences are in addition to possible differences in informational
reliability). Further, speech is claimed to have a privileged status
compared to other sensory stimuli (Edmiston \& Lupyan, 2015; Lupyan \&
Thompson-Schill, 2012; Vouloumanos \& Waxman, 2014; Waxman \& Gelman,
2009; Waxman \& Markow, 1995), and that this privilege is suggested to
be specifically related to the ability to refer (Waxman \& Gelman,
2009).\footnote{There is, however, a debate as to whether speech is
  privileged for children and adults for similar reasons. Whereas some
  researchers suggest that speech is privileged for both children and
  adults because of its ability to refer (e.g., Waxman \& Gelman, 2009),
  others suggest that speech might \emph{not} have a referential status
  from the start. Rather, speech might be preferred by children only
  because of a low level auditory ``overshadowing'' (e.g., Sloutsky \&
  Napolitano, 2003).} Thus, in a referential context, it is possible
that listeners do not treat the auditory and visual modalities as
equivalent sources of information. Instead, there could be a sub-optimal
bias for the auditory modality beyond what is expected from
informational reliability alone.

@@

We investigate how cues from the auditory and the visual modality are
combined in recognizing words in a referential context. In particular,
we study how this combination is performed under various degrees of
uncertainty in both the auditory and the visual modality. Imagine, for
example, that someone is uncertain whether they heard \enquote{pea} or
\enquote{bee.} Does this uncertainty make them rely more on the referent
(e.g., the object being pointed at)? Or, if they are not sure if they
saw a bee or a fly, does this uncertainty make them rely more on the
sound? More importantly, when input in both modalities is uncertain to
varying degrees, do they weight each modality according to its relative
reliability (the optimal strategy), or do they over-rely on a particular
modality?

We begin by proposing an ideal observer model that performs the
combination in an optimal fashion. We then compare the predictions of
the optimal model to human responses. Humans can deviate from the ideal
for several reasons. For instance, as mentioned above, a sub-optimality
can be induced by the privileged status of a particular modality or by
the arbitrariness of the referential association. In order to study
possible patterns of sub-optimality, we compare the optimal model (which
provides a normative benchmark) to a descriptive model (which is fit to
human responses). Comparing parameter estimates between these two
formulations allows us to quantify the degree of deviation from
optimality.

We tested the ideal observer model's predictions in three behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership only. In Experiment 2, we intervened by adding perceptual
noise to the auditory modality, and in Experiment 3, we intervened by
adding perceptual noise to the visual modality. In all experiments,
participants were quantitatively near-optimal, though overall response
precision was slightly lower than expected. In Experiment 1 -- where
neither of the modalities was perturbed with background noise --
participants weighted auditory and visual cues according to the relative
reliability predicted by the optimal model. However, in Experiment 2 and
3, participants over-relied on one modality when the other modality was
perturbed with additional noise.

@@

We introduce a paradigm adapted from a task used by Sloutsky and
Napolitano (2003). The original was used with both children and adults
to probe audio-visual encoding (see Robinson \& Sloutsky, 2010 for
review). Here we use a slightly different version to test word
recognition in a referential context. We use two visual categories (cat
and dog) and two auditory categories (/b/ and /d/ embedded in the
minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is
set between the auditory and the visual categories, leading to two
audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each
trial, participants are presented with an audio-visual target (the
prototype of the target category), immediately followed by an
audio-visual test stimulus (Figure~\ref{fig:task}). The test stimulus
may differ from the target in both the auditory and the visual
components. After these two presentations, participants press
\enquote{same} or \enquote{different.}

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002), which typically studies continuous stimuli (e.g., size,
location), the probabilistic information in our case cannot be
characterized with \emph{sensory noise} only. Since our task involves
responses over categorical variables (phonemes and concepts), the
optimal model should take into account not only the noise variability
around an individual perceptual estimate but also its \emph{categorical
variability}, i.e., the uncertainty related to whether this perceptual
estimate belongs to a given category (see also Bankieris, Bejjanki, \&
Aslin, 2017; Bejjanki et al., 2011). In what follows, we describe a
model that accounts for both types of variability. First, we describe
the model in the simplified case of categorical variability only.
Second, we augment this simplified model to account for sensory noise.

@@

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their crucial parameters (i.e., variances
associated with the visual and auditory modalities) are derived from
data in the unimodal conditions. In addition to these normative models,
we consider a \emph{descriptive} model. It is formally identical to the
normative optimal model (Equation 1), except that the parameters are fit
to actual responses in the bimodal condition. If the referential task
induces sub-optimality (due, for instance, to the arbitrary nature of
the sound-object association), then the descriptive model should explain
more variance than the optimal model does.

@@

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant (but see Experiment 4). Instead, models were constructed
with data collapsed across all participants. The fit was done with a
nonlinear least squares regression using the NLS package in R (Bates \&
Watts, 1988). We computed the values of the parameters, as well as their
95\% confidence intervals, through non-parametric bootstrap (using 10000
iterations).

@@

\caption{\label{tab:param}Statistics for the dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Auditory} & \multicolumn{2}{c}{Visual} & \multicolumn{3}{c}{Bimodal} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5} \cmidrule(l{2pt}r{2pt}){6-8}
Experiment & b\textsubscript{A} & Var\textsubscript{A} & b\textsubscript{V} & Var\textsubscript{V} & b\textsubscript{b} & Var\textsubscript{Ab} & Var\textsubscript{Vb}\\
\midrule
Experiment1 & 0.20 & 2.04 & 0.12 & 3.33 & 0.34 & 4.96 & 7.06\\
Experiment2 & 0.18 & 4.70 & 0.24 & 3.93 & 0.38 & 9.84 & 5.21\\
Experiment3 & 0.24 & 1.94 & -0.11 & 13.00 & 0.35 & 3.00 & 39.42\\
Experiment4 & 0.40 & 1.92 & 0.22 & 3.24 & 0.42 & 4.17 & 7.28\\
\bottomrule
\end{tabular}
\end{table}

@@

This experiment studied the way participants combine multimodal
information to recognize novel words. We found that the optimal model
explained more variance than the auditory or the visual models did,
indicating that particiapants take into account both the auditory and
visual cues when making a decision. That said, Figure \ref{fig:bimodal}
shows that the participants deviated slightly --- but systematically---
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This fact was captured by the
increase in the value of the variance associated with each modality (as
can be noted from Table \ref{tab:param}). Note, however, that despite
this increase response randomness in the bimodal condition, our analysis
of modality preference showed that the \emph{relative} values of these
variances were not different (Figure~\ref{fig:bias}), meaning that there
was no evidence for a modality preference.

@@

We recruited a planned sample of \(N=\) 50 participants from Amazon
Mechanical Turk. Only participants with US IP addresses and a task
approval rate above 99\% were allowed to participate. They were paid at
an hourly rate of \$6/hour. Participants were excluded if they reported
having experienced a technical problem of any sort during the online
experiment (\(N=\) 0), or if they had less than 75\% accurate responses
on the unambiguous training trials (\(N=\) 7). The final sample
consisted of \(N =\) 43 participants. All participants provided informed
consent before taking the
experiment.\footnote{The sample size, exclusion criteria and the main analysese were pre-registered at https://osf.io/h7mzp/.}

@@

The design and procedure were similar to Experiment 1. There were,
however, two differences: 1) We increased the number of responses
elicited per subject from 70 to 300, and 2) we randomized the order of
the three blocks (i.e., visual-only, auditory-only, and audio-visual)
\emph{within} subject: Each participant saw the 3 blocks exactly 6
times, covering all possible ordering combinations. Unlike the
between-subjet randomzation in Experiment 1, this choices allowed us to
avoid any confounding effects due to the order of exposure.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/weights-1} \caption{The histogram shows the distribution of the participants' optimal cue weighing  relative the observed (i.e., descriptive) cue weigting (see Figure 3). Participant are optimal when this relative value is equal to 1 (red solid line). Values larger than 1 suggest over-reliance on the auditory modality whereas values lower than 1 suggest over-reliance on the visual modality. The density plot shows the distribution of simulated individual responses from the population-level probabilistic model. The dashed lines represent 95\% confidence intreval on this simulated distribution. Participants that are outside this interval can be understood as over-relying on the auditory or visual modality in a non-random way (i.e., with p < 0.05).}\label{fig:weights}
\end{figure}

@@

More generally, the idea is that it is easier to integrate correlated,
redundant cues (e.g., the size of a an object based on both visual and
haptic information) into a unified percept, than it is to combine
perceptually uncorrelated cues to make a categorization judgement ,e.g.,
recognizing a freshly learned object category based on its distinctive
features such as shape and color. We argue that our task --- which
consists in combining perceptually uncorrelated cues from the auditory
and visual modalities to recognize a novel word --- is more similar to
the second case.

>>>>

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={How Optimal is Word Recognition Under Multimodal Uncertainty?},
            pdfauthor={Abdellah Fourtassi~\& Michael C. Frank},
            pdfkeywords={Language understanding; audio-visual processing; word learning; speech
perception; computational modeling.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{How Optimal is Word Recognition Under Multimodal Uncertainty?}
    \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C.
Frank\textsuperscript{1}}
    \date{}

@@

\abstract{
Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We study how
adults identify novel words under joint uncertainty in the auditory and
visual modalities and we propose an ideal observer model of how cues in
these modalities are combined optimally. Model predictions are tested in
four experiments where recognition is made under various sources of
uncertainty. We found that participants use both auditory and visual
cues to recognize novel words. When the signal is not distorted with
environmental noise, participants weight the auditory and visual cues
optimally, that is, according to the relative reliability of each
modality. In contrast, when one modality has noise added to it, human
perceivers systematically prefer the unperturbed modality to a greater
extent than the optimal model does. This work extends the literature of
perceptual cue combination to the case of word recognition in a
referential context, thus offering a first step towards a model that
accounts for sound-meaning synergies in early word learning.

@@

Language uses symbols expressed in one modality --- the auditory
modality, in the case of speech --- to communicate about the world,
which we perceive through many different sensory modalities. Consider
hearing someone yell \enquote{bee!} at a picnic, as a honey bee buzzes
around the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence.

However, word identification takes place in a noisy world, and the cues
received through each modality may not provide a definitive answer. On
the auditory side, individual acoustic word tokens are almost always
ambiguous with respect to the particular sequence of phonemes they
represent, which is due to the inherent variability of how a phonetic
category is realized acoustically (Hillenbrand, Getty, Clark, \&
Wheeler, 1995). And some tokens may be distorted additionally by
mispronunciation or ambient noise. Perhaps the speaker was yelling
\enquote{pea} and not \enquote{bee.} Similarly, a sensory impression may
not be enough to make a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead. How does the listener
deal with such multimodal uncertainty to recognize the speaker's
intended word?

As a simplified case study of early word learning, the task of matching
sounds to corresponding visual objects has been studied extensively in
the developmental literature. For example, many studies focus on how
children might succeed in this type of task despite referential
ambiguity (Medina, Snedeker, Trueswell, \& Gleitman, 2011; Pinker, 1989;
Smith \& Yu, 2008; Suanda, Mugwanya, \& Namy, 2014; Vlach \& Johnson,
2013; Vouloumanos, 2008; Yurovsky \& Frank, 2015). However, even when
they have learned the exact meaning of a word, observers (both children
and adults) often still find it challenging to recognize which word the
speaker has uttered, especially under noise (Mattys, Davis, Bradlow, \&
Scott, 2012; Peelle, 2018). The purpose of the current study is thus to
explore word recognition by adults under multimodal uncertainty,
focusing on the special case where people have access to multimodal cues
from the auditory speech and the visual referent. In the General
Discussion, we return to the question of how these findings relate to
questions about word learning.

@@

Some prior ideal observer studies are closely related to the question we
are addressing in the current work. For instance, Clayards et al. (2008)
simulated auditory uncertainty by manipulating the probability
distribution of a cue (Voice Onset Time) that differentiated similar
words (e.g., \enquote{beach} and \enquote{peach}). They found that
humans were sensitive to these probabilistic cues and their judgments
closely reflected the optimal predictions. And Feldman et al. (2009)
studied the perceptual magnet effect, a phenomenon that involves reduced
discriminability near prototypical sounds in the native language (Kuhl,
1991), showing that this effect can be explained as the consequence of
optimally solving the problem of perception under uncertainty (see also
Kronrod, Coppess, \& Feldman, 2016).

Besides the acoustic cues explored in Clayards et al. (2008) and Feldman
et al. (2009), there is extensive evidence that information from the
visual modality, such as the speaker's facial features, also influences
speech understanding (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how probabilistic cues from speech and lip movements
can be optimally combined. They showed that human performance during
audio-visual phonemic labeling was consistent (at least at the
qualitative level) with the predictions of an ideal observer. This
previous research did not, however, study speech understanding when
visual information was obtained through the referential context rather
than through observation of speaker's face. Although some experimental
findings show that information about the identity of a referent can be
integrated with linguistic information to resolve lexical and syntactic
ambiguities in speech (e.g., Eberhard, Spivey-Knowlton, Sedivy, \&
Tanenhaus, 1995; Spivey, Tanenhaus, Eberhard, \& Sedivy, 2002;
Tanenhaus, Spivey-Knowlton, Eberhard, \& Sedivy, 1995), to our knowledge
no study has offered an ideal observer analysis of this task.

Combining information between words and visual referents might seem
similar to audio-visual speech integration (e.g., Bejjanki et al.,
2011), but there are at least two fundamental differences between these
two cases, and both can influence the way the auditory and visual cues
are combined.

First, in the case of audio-visual speech, both modalities offer
information about the same underlying speech category. They differ only
in terms of their informational reliability. In a referential context,
however, the auditory and visual modalities play different roles in the
referential process -- the auditory input represents the \emph{symbol}
whereas the visual input represents the \emph{meaning} (and these
differences are in addition to possible differences in informational
reliability). Further, speech is claimed to have a privileged status
compared to other sensory stimuli (Edmiston \& Lupyan, 2015; Lupyan \&
Thompson-Schill, 2012; Vouloumanos \& Waxman, 2014; Waxman \& Gelman,
2009; Waxman \& Markow, 1995), and this privilege is suggested to be
specifically related to the ability to refer (Waxman \& Gelman,
2009).\footnote{There is, however, a debate as to whether speech is
  privileged for children and adults for similar reasons. Whereas some
  researchers suggest that speech is privileged for both children and
  adults because of its ability to refer (e.g., Waxman \& Gelman, 2009),
  others suggest that speech might \emph{not} have a referential status
  from the start. Rather, speech might be preferred by children only
  because of a low level auditory ``overshadowing'' (e.g., Sloutsky \&
  Napolitano, 2003).} Thus, in a referential context, it is possible
that listeners do not treat the auditory and visual modalities as
equivalent sources of information. Instead, there could be a sub-optimal
bias for the auditory modality beyond what is expected from
informational reliability alone.

@@

We investigate how cues from the auditory and the visual modality are
combined to recognize novel words in a referential context. In
particular, we study how this combination is performed under various
degrees of uncertainty in both the auditory and the visual modality.
Imagine, for example, that someone is uncertain whether they heard
\enquote{pea} or \enquote{bee.} Does this uncertainty make them rely
more on the referent (e.g., the object being pointed at)? Or, if they
are not sure if they saw a bee or a fly, does this uncertainty make them
rely more on the sound? More importantly, when input in both modalities
is uncertain to varying degrees, do they weight each modality according
to its relative reliability (the optimal strategy), or do they over-rely
on a particular modality?

We begin by proposing an ideal observer model that performs the
combination in an optimal fashion. We then compare the predictions of
the optimal model to human responses. Humans can deviate from the ideal
for several reasons. For instance, as mentioned above, a sub-optimality
can be induced by the privileged status of a particular modality or by
the arbitrariness of the referential association. In order to study
possible patterns of sub-optimality, we compare the optimal normative
model to a descriptive model (which is fit to actual responses).
Comparing parameter estimates between these two formulations allows us
to quantify the degree of deviation from optimality.

We tested the ideal observer model's predictions in four behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership (in addition to sensory noise). In Experiment 2, we
intervened by adding environmental noise that degraded information from
the auditory modality and in Experiment 3 we intervened by adding
environmental noise that degraded information from the visual modality.
Finally, Experiment 4 is a replication of Experiment 1 with a higher
power design, allowing us to test cue combination at the individual
level.

@@

We introduce an experimental paradigm adapted from a task used by
Sloutsky and Napolitano (2003). The original was used with both children
and adults to probe audio-visual encoding (see Robinson \& Sloutsky,
2010 for review). Here we use a slightly different version to test word
recognition in a referential context. We use two visual categories (cat
and dog) and two auditory categories (/b/ and /d/ embedded in the
minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is
set between the auditory and the visual categories, leading to two
audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each
trial, participants are presented with an audio-visual target (the
prototype of the target category), immediately followed by an
audio-visual test stimulus (Figure~\ref{fig:task}). The test stimulus
may differ from the target in both the auditory and the visual
components. After these two presentations, participants press
\enquote{same} or \enquote{different.}

@@

We construct an ideal observer model that combines probabilistic
information from the auditory and visual modalities. In contrast to the
model used in most research on multisensory integration (e.g., Ernst \&
Banks, 2002), which typically studies continuous stimuli (e.g., size,
location), the probabilistic information in our case cannot be
characterized with \emph{sensory noise} only. Since our task involves
responses over categorical variables (phonemes and concepts), the
optimal model should take into account not only the noise variability
around an individual perceptual estimate but also its \emph{categorical
variability}, i.e., the uncertainty related to whether this perceptual
estimate belongs to a given category (see also Bankieris, Bejjanki, \&
Aslin, 2017; Bejjanki et al., 2011). In what follows, we describe a
probabilistic model that accounts for both types of variability. First,
we describe the model in the simplified case of categorical variability
only. Second, we augment this simplified model to account for sensory
and envirnmental noise.

@@

The optimal model (as well as the auditory and visual baselines) are
\emph{normative} models. Their predictions are made about human data in
the bimodal condition, but their parameters (i.e., variances associated
with the visual and auditory modalities) are derived from data in the
unimodal conditions. In addition to these normative models, we consider
a \emph{descriptive} model. It is formally identical to the normative
optimal model (Equation 1), except that the parameters are fit to actual
responses in the bimodal condition. If the referential task induces
sub-optimality (due, for instance, to the arbitrary nature of the
sound-object association), then the descriptive model should explain
more variance than the optimal model does.

@@

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant (but see Experiment 4). Instead, models were constructed
with data collapsed across all participants. The fit was done with a
nonlinear least squares regression using the NLS package in R (Bates \&
Watts, 1988). We computed the values of the parameters using
non-parametric bootstrap (with 10000 iterations).

@@

\caption{\label{tab:param}Statistics for the dataset we used.}
\centering
\begin{tabular}[t]{>{\bfseries}lrrrrrrr}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Auditory} & \multicolumn{2}{c}{Visual} & \multicolumn{3}{c}{Bimodal} \\
\cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5} \cmidrule(l{2pt}r{2pt}){6-8}
Experiment & b\textsubscript{A} & Var\textsubscript{A} & b\textsubscript{V} & Var\textsubscript{V} & b\textsubscript{b} & Var\textsubscript{Ab} & Var\textsubscript{Vb}\\
\midrule
Experiment1 & -0.20 & 2.04 & -0.12 & 3.33 & -0.34 & 4.96 & 7.06\\
Experiment2 & -0.18 & 4.70 & -0.24 & 3.93 & -0.38 & 9.84 & 5.21\\
Experiment3 & -0.24 & 1.94 & 0.11 & 13.00 & -0.35 & 3.00 & 39.42\\
Experiment4 & -0.40 & 1.92 & -0.22 & 3.24 & -0.42 & 4.17 & 7.28\\
\bottomrule
\end{tabular}
\end{table}

@@

This experiment studied the way participants combine multimodal
information to recognize novel words. We found that the optimal model
explained more variance than the auditory or the visual models did,
indicating that particiapants take into account both the auditory and
visual cues when making a decision. That said, Figure \ref{fig:bimodal}
shows that the participants deviated slightly --- but systematically---
from the optimal prediction in that they were slightly pulled toward
chance (i.e., the probability 0.5). This fact was captured by the
increase in the value of the variance associated with each modality (as
can be noted from Table \ref{tab:param}). Note, however, that despite
this increase in response randomness, our analysis of modality
preference showed that the \emph{relative} values of these variances
were not different (Figure~\ref{fig:bias}), meaning that there was no
evidence for a modality preference.

@@

We recruited a planned sample of \(N=\) 50 participants from Amazon
Mechanical Turk. Only participants with US IP addresses and a task
approval rate above 99\% were allowed to participate. They were paid at
an hourly rate of \$7/hour. Participants were excluded if they reported
having experienced a technical problem of any sort during the online
experiment (\(N=\) 0), or if they had less than 75\% accurate responses
on the unambiguous training trials (\(N=\) 7). The final sample
consisted of \(N =\) 43 participants. All participants provided informed
consent before taking the
experiment.\footnote{The sample size, exclusion criteria and the main analysese were pre-registered at https://osf.io/h7mzp/.}

@@

The design and procedure were similar to Experiment 1. There were,
however, two differences: 1) We increased the number of responses
elicited per subject from 70 to 300, and 2) we randomized the order of
the three blocks (i.e., visual-only, auditory-only, and audio-visual)
\emph{within} subject: Each participant saw the 3 blocks exactly 6
times, covering all possible ordering combinations. Unlike the
between-subjet randomzation that we used in Experiment 1, this choices
allowed us to avoid a possible confound linked to the order of exposure.

\begin{figure}[!h]
\includegraphics[width=\textwidth]{ms_files/figure-latex/weights-1} \caption{The histogram shows the distribution of the participants' predicted (i.e., optimal) cue weighing relative the observed (i.e., descriptive) weigting (see Figure 3 for the details). The density plot shows the distriubtion of simulated data sampled from the population-level probabilistic model. The dashed lines represent 95\% confidence intreval on this simulated distribution. Optimal beahvior is observed when the value of the relative cue weighing is 1 (red solid line). Participants whose values are outside the confidence interval of the simulated dsitribution over-rely on the visual modalirty (left side) or auditory modality (right side) in a non-random way (i.e., with p < 0.05).}\label{fig:weights}
\end{figure}

@@

Despite this overall near-optomal behavior, we documented two major
cases of sub-optimality. First, in all experiments, the variance
associated with each modality increased in the bimodal condition
compared to the unimodal conditions: Participants responded slightly
more randomly. This increase in randomness could be due limitation on
cognitive resources: Retaining two separate cues at the same time
instead of one cue (as in the unimodal case) is likely to place extra
demands on working memory, causing general performance to drop (see
Mattys \& Wiget, 2011).
